---
title: "ECMP Entry Checks"
author: "Sarah Heidmann"
format: html
editor: visual
date: "2025-09-09"
editor_options: 
  chunk_output_type: inline
---

Please note that this page does not update automatically. It is accurate up to the date above.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse, quietly = TRUE)
library(knitr)
library(DT)

options(knitr.kable.NA = '')
knitr::opts_chunk$set(echo = F)

downdate <- "2025-09-09"
dd <- gsub("-","",downdate)

# Boatlogs
logs_raw <- read_csv(paste0("data/ecdata",dd,"/ECMP_Boatlogs_",downdate,".csv")) %>%
   mutate(Island = ifelse(substr(Fid, 1, 1)==1, "STTSTJ", "STX"),
          ID = paste(Fid, `Date completed`, Team, Role, Name, sep=","))
logs <- logs_raw %>% 
   select(Fid, Island, `Date completed`, Team, Role, Name, ID)

# Abiotic (Plots)
doneplots <- read_csv(paste0("data/ecdata",dd,"/ECMP_Abiotic_all_",downdate,".csv")) %>% 
   mutate(Island = ifelse(substr(Fid, 1, 1)==1, "STTSTJ", "STX"),
          ID = paste(Fid, `Date completed`, Team, Role, Diver, sep=","))
done <- doneplots %>% 
   select(Proofed, Fid, Island, `Date completed`, Team, Role, Diver, ID) %>% 
   unique()

```

# Entry Progress

```{r}
numdone <- length(unique(logs$Fid)) + 300
# % done w fieldwork (600 sites total)
perdonefw <- round(numdone / 600 * 100)
# % done w entry
perdone <- round(length(unique(done$ID)) / length(unique(logs$ID)) * 100)
```

We are `r perdonefw`% done with fieldwork (`r numdone` sites of 600, 300 of which were done in 2024), and `r perdone`% done with entry.

Here is the number of surveys each person has entered:
```{r}
done %>%
   group_by(Diver, Island) %>%
   summarize(n=length(Fid), .groups="drop") %>%
   pivot_wider(names_from = Island, values_from = n) %>% 
   mutate(Total = rowSums(select(.,-Diver), na.rm=TRUE)) %>% 
   arrange(desc(Total)) %>% kable()
```

Out of the entered surveys, here is the habitat breakdown:
```{r}
doneplots %>% 
   group_by(Fid) %>% 
   filter(row_number()==1) %>% 
   group_by(Island, Habitat) %>% 
   summarize(n = length(Fid), .groups="drop") %>% 
   pivot_wider(names_from = Habitat, values_from = n) %>% 
   mutate(Total = rowSums(select(.,-Island), na.rm=TRUE)) %>% 
   kable()
```


# Remaining Data Entry

Here's how many each person has left to enter:

```{r}
todo <- logs[!(logs$ID %in% done$ID),] %>% 
   select(`Date completed`, Island, Fid, Team, Role, Name) %>%
   arrange(`Date completed`, Fid, Team, Role)
todo %>%  
   group_by(Name, Island) %>% 
   summarize(n=length(Fid), .groups= "drop") %>% 
   pivot_wider(names_from = Island, values_from = n) %>% 
   mutate(Total = rowSums(select(.,-Name), na.rm=TRUE)) %>% 
   arrange(desc(Total)) %>% kable()
```

## Full Entry To Do List

```{r}
todo %>% 
   datatable()
```

# QAQC
## Proofing progress

```{r}
totenter <- nrow(done)
totproof <- as.numeric(summary(done$Proofed)[["TRUE"]])
```

Out of `r totenter` surveys entered so far, `r totproof` have been proofed (`r round(totproof/totenter*100)`%).
Here is what each person has left to proof:

```{r}
done %>% 
   filter(!Proofed) %>% 
   group_by(Diver, Island) %>% 
   summarize(n=length(Fid), .groups="drop") %>% 
   pivot_wider(names_from = Island, values_from = n) %>% 
   mutate(Total = rowSums(select(.,-Diver), na.rm=TRUE)) %>% 
   arrange(desc(Total)) %>% kable()
```

### Check for mismatched entries

Are there any surveys that don't match a boatlog entry?

```{r}
surverror <- done %>% 
   filter(!(ID %in% logs$ID))
if(nrow(surverror)>0){
   surverror %>% select(-ID) %>% kable()
} else {
   print("No")
}
```

### Check for duplicates

Are there any duplicated boatlogs?

```{r}
log_dups <- logs[duplicated(logs$ID),]
   
if(nrow(log_dups)>0){
   log_dups %>% select(-ID) %>% kable()
} else {
   print("No")
}
```

Are there any duplicated survey entries?

```{r}
sitediverlist <- done %>% 
   mutate(sitediver = paste(Fid, Diver, sep=","))
done_dups <- sitediverlist[duplicated(sitediverlist$ID),]
if(nrow(done_dups)>0){
   done_dups %>% select(-ID) %>% kable()
} else {
   print("No")
}
```

### Check for habitat agreement
Do any of the habitats disagree between buddies?
```{r}
habsame <- doneplots %>% 
   select(Fid, Role, Habitat) %>% 
   unique() %>% 
   pivot_wider(names_from = Role, values_from = Habitat) %>% 
   mutate(same = A==B) %>% 
   filter(same==FALSE)
if(nrow(habsame)>0){
   habsame %>% select(-same) %>% kable()
} else {
   print("No")
}
```

