---
title: "TCRMP_2023_Benthic_QAQC.qmd"
author: "Sarah Heidmann"
format: html
editor: visual
---

### Sarah Heidmann

### 2025-01-22

QAQC for TCRMP coral health and algae height data for 2023 and 2024 Post Bleaching (PBL). Suggestions for additions and improvement are welcomed.

Luckily, the data entry system catches a lot of easy mistakes like mispellings. We just have to go through the mistakes it might not catch. When we catch some that are fixable, change online and export again, or if the online database has already been cleared, change directly in the data.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE)
library(tidyverse, quietly = TRUE)
library(knitr)
options(knitr.kable.NA = '')
library(DT)

# Read qaqc'd data for fish and diadema
chs_raw_23 <- read_csv("data/2023/TCRMP_CoralHealth_2023_webexport.csv") %>% 
   replace_na(list(`Old mortality`=0, `Recent mortality`=0,
                  `Bl sp`=0,`Bl p`=0, `Bl vp`=0, `Bl bl`=0))
chs_raw_24 <- read_csv("data/2023/TCRMP_CoralHealth_2024PBL_webexport.csv") %>% 
   replace_na(list(`Old mortality`=0, `Recent mortality`=0,
                  `Bl sp`=0,`Bl p`=0, `Bl vp`=0, `Bl bl`=0))
ahs_raw <- read_csv("data/2023/TCRMP_AlgaeHeights_2023_webexport.csv")

# Species master for QAQC numbers
#species <- read_csv("data/fish_species_list_2022.csv")

```

# Overview

How big are the datasets?

```{r}
corals23 <- length(unique(chs_raw_23$`Transect coral`))
corals24 <- length(unique(chs_raw_24$`Transect coral`))
```

Coral health 2023 has `r corals23` corals. Coral health 2024PBL has `r corals24` corals.

Do we have all the coral transects in 2023? (should have 6 transects each at 33 sites)

```{r}
# Summarize number of transects by site
chs_sitesum_23 <- chs_raw_23 %>% 
   group_by(`Site name`,Rep) %>% 
   summarize(n=1, .groups="drop_last") %>% 
   summarize(Transects = sum(n))
# High level look at number of reps
chs_sitesum_23 %>% 
   group_by(Transects) %>% 
   summarize(Sites = length(`Site name`)) %>% kable()
```

Those with less than 6 transects are below. We know we missed one transect at Hind.

```{r}
# See the sitewise list
chs_sitesum_23 %>% filter(Transects<6) %>% arrange(Transects) %>% datatable()
```

Do we have all the coral transects in 2024PBL? (should have 6 transects each at 33 sites)

```{r}
# Summarize number of transects by site
chs_sitesum_24 <- chs_raw_24 %>% 
   group_by(`Site name`,Rep) %>% 
   summarize(n=1, .groups="drop_last") %>% 
   summarize(Transects = sum(n))
# High level look at number of reps
chs_sitesum_24 %>% 
   group_by(Transects) %>% 
   summarize(Sites = length(`Site name`)) %>% kable()
```

Those with less than 6 transects are below.

```{r}
# See the sitewise list
chs_sitesum_24 %>% filter(Transects<6) %>% arrange(Transects) %>% datatable()
```

Do we have all the algae transects in 2023? (should have 6 transects each at 33 sites)

```{r}
# Summarize number of transects by site
ahs_sitesum <- ahs_raw %>% 
   group_by(`Site name`,Rep) %>% 
   summarize(n=1, .groups="drop_last") %>% 
   summarize(Transects = sum(n))
# High level look at number of reps
ahs_sitesum %>% 
   group_by(Transects) %>% 
   summarize(Sites = length(`Site name`)) %>% kable()
```

Those with less than 6 transects are below.

```{r}
# See the sitewise list
ahs_sitesum %>% filter(Transects<6) %>% arrange(Transects) %>% datatable()
```

## Date ranges

```{r}
# Date ranges for each survey type
ch_range_23 <- range(chs_raw_23$`Date completed`)
ch_range_24 <- range(chs_raw_24$`Date completed`)
ah_range <- range(ahs_raw$`Date completed`)
```

Coral health 2023 took place between dates: `r ch_range_23`

Coral health 2024PBL took place between dates: `r ch_range_24`

Algae heights took place between dates: `r ah_range`

## Notes

Are there any notes we should know about?

```{r}
chs_raw_23 %>% filter(!is.na(`Site Notes`)) %>% 
   pull(`Site Notes`) %>% 
   unique()
chs_raw_24 %>% filter(!is.na(`Site Notes`)) %>% 
   pull(`Site Notes`) %>% 
   unique()
ahs_raw %>% filter(!is.na(Notes)) %>% 
   pull(Notes) %>% 
   unique()
```

# Corals

## Occurrence

How often did we see each coral species, overall and by diver? You can use this table to look at rare species and possible mis-IDs.

In 2023:

```{r}
chs_corals_23 <- chs_raw_23 %>% 
   select(`Site name`, Rep, Name, Method, `Transect coral`, 
          `Coral full name`, `Length cm`, `Width cm`, `Height cm`,
          `Bl sp`, `Bl p`, `Bl vp`, `Bl bl`, 
          `Old mortality`, `Recent mortality`) %>% 
   unique()
chs_corals_23 %>% 
   group_by(`Coral full name`, Name) %>% 
   summarize(n = length(`Transect coral`), .groups = "drop") %>% 
   pivot_wider(id_cols = c(`Coral full name`), 
               names_from = Name, values_from = n, values_fill = 0) %>% 
   mutate(Total = rowSums(select(.,-`Coral full name`)), 
          .after = `Coral full name`) %>% 
   datatable()
chs_corals_23 %>% 
   ggplot(data=., aes(x=`Coral full name`, fill = Name)) +
   geom_bar(aes(y = after_stat(count)/sum(after_stat(count)))) +
   ylab("Proportion of Corals Seen")+
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

In 2024 PBL:

```{r}
chs_corals_24 <- chs_raw_24 %>% 
   select(`Site name`, Rep, Name, Method, `Transect coral`,
          `Coral full name`, `Length cm`, `Width cm`, `Height cm`,
          `Bl sp`, `Bl p`, `Bl vp`, `Bl bl`, 
          `Old mortality`, `Recent mortality`) %>% 
   unique() 
chs_corals_24 %>% 
   group_by(`Coral full name`, Name) %>% 
   summarize(n = length(`Transect coral`), .groups = "drop") %>% 
   pivot_wider(id_cols = c(`Coral full name`), 
               names_from = Name, values_from = n, values_fill = 0) %>% 
   mutate(Total = rowSums(select(.,-`Coral full name`)), 
          .after = `Coral full name`) %>% 
   datatable()
chs_corals_24 %>% 
   ggplot(data=., aes(x=`Coral full name`, fill = Name)) +
   geom_bar(aes(y = after_stat(count)/sum(after_stat(count)))) +
   ylab("Proportion of Corals Seen")+
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

## Abundance

How many corals did we see on average on transects across sites?

In 2023:

```{r}
 chs_corals_23 %>% 
   group_by(`Site name`, Method, Rep) %>% 
   summarize(Corals = length(`Transect coral`), .groups = "drop_last") %>% 
   summarize(mean = round(mean(Corals),1), sd = round(sd(Corals),1),
             min = min(Corals), max = max(Corals),
             .groups="drop") %>% 
   arrange(desc(mean)) %>% datatable()
```

In 2024 PBL:

```{r}
chs_corals_24 %>% 
   group_by(`Site name`, Method, Rep) %>% 
   summarize(Corals = length(`Transect coral`), .groups = "drop_last") %>% 
   summarize(mean = round(mean(Corals),1), sd = round(sd(Corals),1),
             min = min(Corals), max = max(Corals),
             .groups="drop") %>% 
   arrange(desc(mean)) %>% datatable()
```

## Sizing

What did the max dimension (length in cm) of each coral species look like?

In 2023:

```{r}
# Do we want to add width and height to this check?
# See Fig 4 of ncrmp for example.
# would need to pivot_longer the dimensions then summarize then wrap.
chs_corals_23 %>% 
   group_by(`Coral full name`) %>% 
   summarize(n = length(`Length cm`),
             mean = round(mean(`Length cm`),1), sd = round(sd(`Length cm`),1),
             min = min(`Length cm`), max = max(`Length cm`),
             .groups = "drop") %>% 
   arrange(`Coral full name`) %>% datatable()
```

In 2024 PBL:

```{r}
chs_corals_24 %>% 
   group_by(`Coral full name`) %>% 
   summarize(n = length(`Length cm`),
             mean = round(mean(`Length cm`),1), sd = round(sd(`Length cm`),1),
             min = min(`Length cm`), max = max(`Length cm`),
             .groups = "drop") %>% 
   arrange(`Coral full name`) %>% datatable()
```

Let's look at the size frequency distributions of the top 10 most common species.

In 2023:

```{r}
commcoral <- chs_corals_23 %>% 
   group_by(`Coral full name`) %>% 
   summarize(n = length(`Coral full name`)) %>% 
   arrange(desc(n)) %>% slice_head(n=10) %>% pull(`Coral full name`)
chs_corals_23 %>% 
   filter(`Coral full name` %in% commcoral) %>% 
   ggplot(data=.) + 
   geom_histogram(aes(x=`Length cm`), binwidth=10) + 
   facet_wrap(~`Coral full name`, scales = "free")
```

In 2024PBL:

```{r}
chs_corals_24 %>% 
   filter(`Coral full name` %in% commcoral) %>% 
   ggplot(data=.) + 
   geom_histogram(aes(x=`Length cm`), binwidth=10) + 
   facet_wrap(~`Coral full name`, scales = "free")
```

## Mortality

What is the prevalence of old and new mortality by coral species?

In 2023:

```{r}
chs_corals_23 %>%  
   mutate(`Old mortality`= ifelse(`Old mortality`>0,1,0),
          `Recent mortality`= ifelse(`Recent mortality`>0,1,0)) %>% 
   group_by(`Coral full name`) %>% 
   summarize(`Old mortality`=sum(`Old mortality`)/length(`Old mortality`),
             `Recent mortality`=sum(`Recent mortality`)/length(`Recent mortality`),
             .groups = "drop") %>% 
   pivot_longer(cols = c(`Old mortality`, `Recent mortality`), 
                names_to = "Type", values_to = "value") %>% 
   ggplot(data=., aes(x=`Coral full name`, fill =Type)) +
   geom_bar(stat = "identity", position = "dodge", 
            aes(y = value)) +
   #geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

In 2024 PBL:

```{r}
chs_corals_24 %>%  
   mutate(`Old mortality`= ifelse(`Old mortality`>0,1,0),
          `Recent mortality`= ifelse(`Recent mortality`>0,1,0)) %>% 
   group_by(`Coral full name`) %>% 
   summarize(`Old mortality`=sum(`Old mortality`)/length(`Old mortality`),
             `Recent mortality`=sum(`Recent mortality`)/length(`Recent mortality`),
             .groups = "drop") %>% 
   pivot_longer(cols = c(`Old mortality`, `Recent mortality`), 
                names_to = "Type", values_to = "value") %>% 
   ggplot(data=., aes(x=`Coral full name`, fill =Type)) +
   geom_bar(stat = "identity", position = "dodge", 
            aes(y = value)) +
   #geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

What is the prevalence of old and new mortality by diver? Check for possible over- or under-estimating.

In 2023:

```{r}
chs_corals_23 %>%  
   mutate(`Old mortality`= ifelse(`Old mortality`>0,1,0),
          `Recent mortality`= ifelse(`Recent mortality`>0,1,0)) %>% 
   group_by(Name) %>% 
   summarize(`Old mortality`=sum(`Old mortality`)/length(`Old mortality`),
             `Recent mortality`=sum(`Recent mortality`)/length(`Recent mortality`),
             .groups = "drop") %>% 
   pivot_longer(cols = c(`Old mortality`, `Recent mortality`), 
                names_to = "Type", values_to = "value") %>% 
   add_row(Name = "Overall", Type="Old mortality", 
           value = nrow(filter(chs_corals_23,`Old mortality`>0))/nrow(chs_corals_23)) %>% 
   add_row(Name = "Overall", Type="Recent mortality", 
           value = nrow(filter(chs_corals_23,`Recent mortality`>0))/nrow(chs_corals_23)) %>% 
   ggplot(data=., aes(x=Name, fill =Type)) +
   geom_bar(stat = "identity", position = "dodge", 
            aes(y = value)) +
   #geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

In 2024PBL:

```{r}
chs_corals_24 %>%  
   mutate(`Old mortality`= ifelse(`Old mortality`>0,1,0),
          `Recent mortality`= ifelse(`Recent mortality`>0,1,0)) %>% 
   group_by(Name) %>% 
   summarize(`Old mortality`=sum(`Old mortality`)/length(`Old mortality`),
             `Recent mortality`=sum(`Recent mortality`)/length(`Recent mortality`),
             .groups = "drop") %>% 
   pivot_longer(cols = c(`Old mortality`, `Recent mortality`), 
                names_to = "Type", values_to = "value") %>% 
   add_row(Name = "Overall", Type="Old mortality", 
           value = nrow(filter(chs_corals_24,`Old mortality`>0))/nrow(chs_corals_24)) %>% 
   add_row(Name = "Overall", Type="Recent mortality", 
           value = nrow(filter(chs_corals_24,`Recent mortality`>0))/nrow(chs_corals_24)) %>% 
   ggplot(data=., aes(x=Name, fill =Type)) +
   geom_bar(stat = "identity", position = "dodge", 
            aes(y = value)) +
   #geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

What is the extent of old and new mortality by coral species?

In 2023:

```{r, warning=FALSE}
chs_corals_23 %>%  
   pivot_longer(cols = c(`Old mortality`, `Recent mortality`), 
                names_to = "Type", values_to = "value") %>% 
   group_by(`Coral full name`, Type) %>% 
   summarize(mean = round(mean(value),2), sd = round(sd(value),2), 
             n = length(value), 
             .groups = "drop") %>% 
   ggplot(data=., aes(x=`Coral full name`, fill =Type)) +
   geom_bar(stat = "identity", position = "stack", 
            aes(y = mean)) +
   #geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

In 2024 PBL:

```{r, warning=FALSE}
chs_corals_24 %>%  
   pivot_longer(cols = c(`Old mortality`, `Recent mortality`), 
                names_to = "Type", values_to = "value") %>% 
   group_by(`Coral full name`, Type) %>% 
   summarize(mean = round(mean(value),2), sd = round(sd(value),2), 
             n = length(value), 
             .groups = "drop") %>% 
   ggplot(data=., aes(x=`Coral full name`, fill =Type)) +
   geom_bar(stat = "identity", position = "stack", 
            aes(y = mean)) +
   #geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

What is the extent of old and new mortality by diver? Check for possible over- or under-estimating.

In 2023:

```{r}
chs_corals_23 %>%  
   pivot_longer(cols = c(`Old mortality`, `Recent mortality`), 
                names_to = "Type", values_to = "value") %>% 
   group_by(Name, Type) %>% 
   summarize(mean = round(mean(value),2), sd = round(sd(value),2), 
             n = length(value), 
             .groups = "drop") %>% 
   add_row(Name = "Overall", Type="Old mortality", 
           mean = round(mean(chs_corals_23$`Old mortality`),2), 
           sd = round(sd(chs_corals_23$`Old mortality`),2),
           n = length(chs_corals_23$`Old mortality`)) %>% 
   add_row(Name = "Overall", Type="Recent mortality", 
           mean = round(mean(chs_corals_23$`Recent mortality`),2), 
           sd = round(sd(chs_corals_23$`Recent mortality`),2),
           n = length(chs_corals_23$`Recent mortality`)) %>% 
   ggplot(data=., aes(x=Name, fill =Type)) +
   geom_bar(stat = "identity", position = "stack", 
            aes(y = mean)) +
   #geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

In 2024 PBL:

```{r}
chs_corals_24 %>%  
   pivot_longer(cols = c(`Old mortality`, `Recent mortality`), 
                names_to = "Type", values_to = "value") %>% 
   group_by(Name, Type) %>% 
   summarize(mean = round(mean(value),2), sd = round(sd(value),2), 
             n = length(value), 
             .groups = "drop") %>% 
   add_row(Name = "Overall", Type="Old mortality", 
           mean = round(mean(chs_corals_24$`Old mortality`),2), 
           sd = round(sd(chs_corals_24$`Old mortality`),2),
           n = length(chs_corals_24$`Old mortality`)) %>% 
   add_row(Name = "Overall", Type="Recent mortality", 
           mean = round(mean(chs_corals_24$`Recent mortality`),2), 
           sd = round(sd(chs_corals_24$`Recent mortality`),2),
           n = length(chs_corals_24$`Recent mortality`)) %>% 
   ggplot(data=., aes(x=Name, fill =Type)) +
   geom_bar(stat = "identity", position = "stack", 
            aes(y = mean)) +
   #geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

## Bleaching

What is the prevalence of bleaching types by coral species?

In 2023:

```{r}
chs_corals_23 %>%  
   mutate(`Bl sp`= ifelse(`Bl sp`>0,1,0),
          `Bl p`= ifelse(`Bl p`>0,1,0),
          `Bl vp`= ifelse(`Bl vp`>0,1,0),
          `Bl bl`= ifelse(`Bl bl`>0,1,0)) %>% 
   group_by(`Coral full name`) %>% 
   summarize(`Bl sp`=sum(`Bl sp`)/length(`Bl sp`),
             `Bl p`=sum(`Bl p`)/length(`Bl p`),
             `Bl vp`=sum(`Bl vp`)/length(`Bl vp`),
             `Bl bl`=sum(`Bl bl`)/length(`Bl bl`),
             .groups = "drop") %>% 
   pivot_longer(!`Coral full name`, 
                names_to = "Type", values_to = "value") %>% 
   ggplot(data=., aes(x=`Coral full name`, fill =Type)) +
   geom_bar(stat = "identity", position = "dodge", 
            aes(y = value)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

In 2024 PBL:

```{r}
chs_corals_24 %>%  
   mutate(`Bl sp`= ifelse(`Bl sp`>0,1,0),
          `Bl p`= ifelse(`Bl p`>0,1,0),
          `Bl vp`= ifelse(`Bl vp`>0,1,0),
          `Bl bl`= ifelse(`Bl bl`>0,1,0)) %>% 
   group_by(`Coral full name`) %>% 
   summarize(`Bl sp`=sum(`Bl sp`)/length(`Bl sp`),
             `Bl p`=sum(`Bl p`)/length(`Bl p`),
             `Bl vp`=sum(`Bl vp`)/length(`Bl vp`),
             `Bl bl`=sum(`Bl bl`)/length(`Bl bl`),
             .groups = "drop") %>% 
   pivot_longer(!`Coral full name`, 
                names_to = "Type", values_to = "value") %>% 
   ggplot(data=., aes(x=`Coral full name`, fill =Type)) +
   geom_bar(stat = "identity", position = "dodge", 
            aes(y = value)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

What is the prevalence of bleaching types by diver? Check for possible over- or under-estimating.

In 2023:

```{r}
chs_corals_23 %>%  
   mutate(`Bl sp`= ifelse(`Bl sp`>0,1,0),
          `Bl p`= ifelse(`Bl p`>0,1,0),
          `Bl vp`= ifelse(`Bl vp`>0,1,0),
          `Bl bl`= ifelse(`Bl bl`>0,1,0)) %>% 
   group_by(Name) %>% 
   summarize(`Bl sp`=sum(`Bl sp`)/length(`Bl sp`),
             `Bl p`=sum(`Bl p`)/length(`Bl p`),
             `Bl vp`=sum(`Bl vp`)/length(`Bl vp`),
             `Bl bl`=sum(`Bl bl`)/length(`Bl bl`),
             .groups = "drop") %>% 
   pivot_longer(!Name, 
                names_to = "Type", values_to = "value") %>% 
   add_row(Name = "Overall", Type="Bl sp", 
           value = nrow(filter(chs_corals_23,`Bl sp`>0))/nrow(chs_corals_23)) %>% 
   add_row(Name = "Overall", Type="Bl p", 
           value = nrow(filter(chs_corals_23,`Bl p`>0))/nrow(chs_corals_23)) %>% 
   add_row(Name = "Overall", Type="Bl vp", 
           value = nrow(filter(chs_corals_23,`Bl vp`>0))/nrow(chs_corals_23)) %>% 
   add_row(Name = "Overall", Type="Bl bl", 
           value = nrow(filter(chs_corals_23,`Bl bl`>0))/nrow(chs_corals_23)) %>% 
   ggplot(data=., aes(x=Name, fill =Type)) +
   geom_bar(stat = "identity", position = "dodge", 
            aes(y = value)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

In 2024PBL:

```{r}
chs_corals_24 %>%  
   mutate(`Bl sp`= ifelse(`Bl sp`>0,1,0),
          `Bl p`= ifelse(`Bl p`>0,1,0),
          `Bl vp`= ifelse(`Bl vp`>0,1,0),
          `Bl bl`= ifelse(`Bl bl`>0,1,0)) %>% 
   group_by(Name) %>% 
   summarize(`Bl sp`=sum(`Bl sp`)/length(`Bl sp`),
             `Bl p`=sum(`Bl p`)/length(`Bl p`),
             `Bl vp`=sum(`Bl vp`)/length(`Bl vp`),
             `Bl bl`=sum(`Bl bl`)/length(`Bl bl`),
             .groups = "drop") %>% 
   pivot_longer(!Name, 
                names_to = "Type", values_to = "value") %>% 
   add_row(Name = "Overall", Type="Bl sp", 
           value = nrow(filter(chs_corals_24,`Bl sp`>0))/nrow(chs_corals_23)) %>% 
   add_row(Name = "Overall", Type="Bl p", 
           value = nrow(filter(chs_corals_24,`Bl p`>0))/nrow(chs_corals_23)) %>% 
   add_row(Name = "Overall", Type="Bl vp", 
           value = nrow(filter(chs_corals_24,`Bl vp`>0))/nrow(chs_corals_23)) %>% 
   add_row(Name = "Overall", Type="Bl bl", 
           value = nrow(filter(chs_corals_24,`Bl bl`>0))/nrow(chs_corals_23)) %>% 
   ggplot(data=., aes(x=Name, fill =Type)) +
   geom_bar(stat = "identity", position = "dodge", 
            aes(y = value)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

What is the extent of bleaching types by coral species?

In 2023:

```{r, warning=FALSE}
chs_corals_23 %>%  
   pivot_longer(cols = starts_with("Bl"), 
                names_to = "Type", values_to = "value") %>% 
   group_by(`Coral full name`, Type) %>% 
   summarize(mean = round(mean(value),2), sd = round(sd(value),2), 
             n = length(value), 
             .groups = "drop") %>% 
   ggplot(data=., aes(x=`Coral full name`, fill =Type)) +
   geom_bar(stat = "identity", position = "stack", 
            aes(y = mean)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

In 2024 PBL:

```{r, warning=FALSE}
chs_corals_24 %>%  
   pivot_longer(cols = starts_with("Bl"),
                names_to = "Type", values_to = "value") %>% 
   group_by(`Coral full name`, Type) %>% 
   summarize(mean = round(mean(value),2), sd = round(sd(value),2), 
             n = length(value), 
             .groups = "drop") %>% 
   ggplot(data=., aes(x=`Coral full name`, fill =Type)) +
   geom_bar(stat = "identity", position = "stack", 
            aes(y = mean)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

What is the extent of bleaching types by diver? Check for possible over- or under-estimating.

In 2023:

```{r}
chs_corals_23 %>%  
   pivot_longer(cols = starts_with("Bl"), 
                names_to = "Type", values_to = "value") %>% 
   group_by(Name, Type) %>% 
   summarize(mean = round(mean(value),2), sd = round(sd(value),2), 
             n = length(value), 
             .groups = "drop") %>% 
   add_row(Name = "Overall", Type="Bl sp", 
           mean = round(mean(chs_corals_23$`Bl sp`),2)) %>% 
   add_row(Name = "Overall", Type="Bl p", 
           mean = round(mean(chs_corals_23$`Bl p`),2)) %>% 
   add_row(Name = "Overall", Type="Bl vp", 
           mean = round(mean(chs_corals_23$`Bl vp`),2)) %>% 
   add_row(Name = "Overall", Type="Bl bl", 
           mean = round(mean(chs_corals_23$`Bl bl`),2)) %>% 
   ggplot(data=., aes(x=Name, fill =Type)) +
   geom_bar(stat = "identity", position = "stack", 
            aes(y = mean)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

In 2024 PBL:

```{r}
chs_corals_24 %>%  
   pivot_longer(cols = starts_with("Bl"), 
                names_to = "Type", values_to = "value") %>% 
   group_by(Name, Type) %>% 
   summarize(mean = round(mean(value),2), sd = round(sd(value),2), 
             n = length(value), 
             .groups = "drop") %>% 
   add_row(Name = "Overall", Type="Bl sp", 
           mean = round(mean(chs_corals_24$`Bl sp`),2)) %>% 
   add_row(Name = "Overall", Type="Bl p", 
           mean = round(mean(chs_corals_24$`Bl p`),2)) %>% 
   add_row(Name = "Overall", Type="Bl vp", 
           mean = round(mean(chs_corals_24$`Bl vp`),2)) %>% 
   add_row(Name = "Overall", Type="Bl bl", 
           mean = round(mean(chs_corals_24$`Bl bl`),2)) %>% 
   ggplot(data=., aes(x=Name, fill =Type)) +
   geom_bar(stat = "identity", position = "stack", 
            aes(y = mean)) +
   theme(axis.text.x = element_text(angle = 45, hjust=1))
```

## Interactions

What were the most common interactions and their average values?

```{r}
chs_raw_23 %>% 
   filter(!is.na(`Interaction name`)) %>% 
   group_by(`Interaction name`) %>% 
   summarize(n = length(`Interaction value`),
             mean = round(mean(`Interaction value`),2),
             sd = round(sd(`Interaction value`),2),
             max = max(`Interaction value`),
             min = min(`Interaction value`)) %>% 
   arrange(desc(n)) %>% datatable()
```

```{r}
# More ideas
# 
# focus on diver trends to generate info for feedback on training.
# 
# coral size flags table. what is our 99% vs ncrmp? how many are flagged?
```

# Algae Heights

## Abundance

How many algal points did we get per transect on average across sites?

```{r}
ahs_raw %>% 
   group_by(`Site name`, Rep) %>% 
   summarize(Algaes = length(`Code name`), .groups = "drop_last") %>% 
   summarize(n=length(Algaes), mean = round(mean(Algaes),1), 
             sd = round(sd(Algaes),1), 
             min = min(Algaes), max = max(Algaes), .groups="drop") %>% 
   arrange(desc(mean)) %>% datatable()
```

Buck STX was 90% CYAN, which was not recorded. We may want to delete this data altogether.

## Sizing

How tall (cm) was each species of algae? You can use this table to look at rare species as well.

```{r}
ahs_raw %>% 
   group_by(`Code name`) %>% 
   summarize(n=length(`Height cm`), mean = round(mean(`Height cm`),1), 
             sd = round(sd(`Height cm`),1), 
             min = min(`Height cm`), max = max(`Height cm`), .groups="drop") %>% 
   arrange(desc(mean)) %>% datatable()
```

FILSED should not be 7.5 cm. What happened there? Let's look at the FILSED \> 2 cm.

```{r}
ahs_raw %>% 
   filter(`Code name`=="FILSED") %>% 
   filter(`Height cm`>2) %>% 
   arrange(desc(`Height cm`)) %>% 
   select(-Notes,-`Algae height`,-`Full name`,-Proofed) %>% datatable()
```

Also 9.6 cm LOBO? They can be quite fleshy at deep sites but that's a lot. Let's look at LOBO \> 3 cm.

```{r}
ahs_raw %>% 
   filter(`Code name`=="LOBO") %>% 
   filter(`Height cm`>3) %>% 
   arrange(desc(`Height cm`)) %>% 
   select(-Notes,-`Algae height`,-`Full name`,-Proofed) %>% datatable()
```

Could these transects be in mm instead of cm? Look at Erin's Magens data and Shaun's BIT data.

```{r}
ahs_raw %>% 
   filter(`Site name`=="Magens Bay", Name == "HOLLANDER_ERIN") %>% 
   select(-Notes,-`Algae height`,-`Full name`,-Proofed) %>% datatable()
ahs_raw %>% 
   filter(`Site name`=="Buck Island STT", Name == "KADISON_SHAUN") %>% 
   select(-Notes,-`Algae height`,-`Full name`,-Proofed) %>% datatable()
```

Erin's look like they might be mm. Shaun has enough decimals that I think they are cm.
