---
title: "TCRMP 2024 Fish QAQC"
author: "Sarah Heidmann"
format: html
editor: visual
toc: true
---

### 2025-03-06

QAQC for TCRMP fish data for 2024. Suggestions for additions and improvement are welcomed.

Luckily, the data entry system catches a lot of easy mistakes like mispellings. We just have to go through the mistakes it might not catch. When we catch some that are fixable, change online and export again, or if the online database has already been cleared, change directly in the data.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE)
library(tidyverse, quietly = TRUE)
library(knitr)
options(knitr.kable.NA = '')
library(DT)

# Read qaqc'd data for fish and diadema
fts_raw <- read_csv("data/2024/TCRMP_FishTransects_2024_webexport.csv") %>% 
   mutate(Total = rowSums(select(.,X0to5:Xgt150), na.rm=T))
frs_raw <- read_csv("data/2024/TCRMP_FishRovers_2024_webexport.csv")
das_raw <- read_csv("data/2024/TCRMP_Diadema_2024_qaqc.csv",
                    col_types=cols(`Date completed`=col_date(format="%m/%d/%Y")))

# Species master for QAQC numbers
species <- read_csv("data/fishlists/website.csv")

```

# Overview

How big are the datasets?

Transects have `r nrow(fts_raw)` transect-species. Rovers have `r nrow(frs_raw)` rover-species.

```{r, tab.cap="Table 1. Do we have all the fish transects? We should have 9 transects each at 33 sites and maybe 1 at Ginsburg's Fringe."}
# Summarize number of transects by site
fts_sitesum <- fts_raw %>% 
   group_by(`Site name`,Rep) %>% 
   summarize(n=1, .groups="drop_last") %>% 
   summarize(Transects = sum(n))
# High level look at number of reps
fts_sitesum %>% 
   group_by(Transects) %>% 
   summarize(Sites = length(`Site name`)) %>% kable()
```

```{r, tab.cap="Table 2. Sites with less than 9 fish transects."}
kable(x=tibble(" "=" ")) # hack caption above datatable
# See the sitewise list
fts_sitesum %>% filter(Transects<9) %>% arrange(Transects) %>% datatable()
```

```{r, tab.cap="Table 3. Do we have all the fish rovers? We should have 3 reps each at 33 sites and maybe 1 at Ginsburg's Fringe."}
# Summarize number of rovers by site
frs_sitesum <- frs_raw %>% 
   group_by(`Site name`,Rep) %>% 
   summarize(n=1, .groups="drop_last") %>% 
   summarize(Rovers = sum(n))
# High level look at number of reps
frs_sitesum %>% 
   group_by(Rovers) %>% 
   summarize(Sites = length(`Site name`)) %>% kable()
```

```{r, tab.cap="Table 4. Sites with less than 3 fish rovers."}
# See the sitewise list
frs_sitesum %>% filter(Rovers <3) %>% arrange(Rovers) %>% datatable()
```

## Date ranges

```{r}
# Date ranges for each survey type
ft_range <- range(fts_raw$`Date completed`)
da_range <- range(das_raw$`Date completed`)
fr_range <- range(frs_raw$`Date completed`)
```

Fish transects took place between dates: `r ft_range`

Diademas took place between dates: `r da_range`

Fish rovers took place between dates: `r fr_range`

## Notes

Are there any notes we should know about?

```{r}
fts_raw %>% filter(!is.na(Notes)) %>% 
   pull(Notes) %>% 
   unique()
frs_raw %>% filter(!is.na(Notes)) %>% 
   pull(Notes) %>% 
   unique()
```

# Abundance

An observation is flagged if the number of fish observed is greater than the maximum number guideline.

## Transect Abundance

```{r}
# Create flags based on master species list
fts_abund <- fts_raw %>% 
   left_join(species,by=c("Scientific name"="scientific name")) %>% 
   mutate(flag = as.factor(ifelse(Total > `max abund`, "yes","no")))
# Calculate percent flagged
tperabundflag <- round(summary(fts_abund$flag)[2] / nrow(fts_abund) *100, 1) 
# Make table of flags
tabund_flag <- fts_abund %>% filter(flag == "yes") %>%
   select(`Site name`,Rep,Name,
          `Scientific name`,`Common name`,Total,`max abund`) %>% 
   mutate(Diff = Total - `max abund`) %>% 
   mutate(RelDiff = round(Diff / `max abund`,1))
```

There were `r tperabundflag`% of transect species flagged on abundance.

```{r, tab.cap="Table 5. Observations from fish transects flagged on high abundance."}
kable(x=tibble(" "=" ")) # hack caption above datatable
datatable(tabund_flag)
```

```{r, tab.cap="Table 6. Fish species that were most frequently flagged on abundance in transects. We might want to change the max number on these."}
kable(x=tibble(" "=" ")) # hack caption above datatable
# Summarize abundance flag table by species
tabund_flag %>%
   group_by(`Common name`,`max abund`) %>%
   summarize(Flags = length(Total), .groups="drop") %>%
   arrange(desc(Flags)) %>% datatable()
```

```{r, tab.cap="Table 7. Divers who were most frequently flagged on high abundance in transects. They might be overestimating counts."}
kable(x=tibble(" "=" ")) # hack caption above datatable
# Summarize total number of surveys for each person
fts_namecount <- fts_raw %>% 
   group_by(Name) %>%
   summarize(Observations = length(Name)) %>% 
   arrange(desc(Observations))
# Standardize flags to a percentage of total fish seen
tabund_flag %>% group_by(Name) %>%
   summarize(Flags = length(Total)) %>%
   left_join(fts_namecount,by="Name") %>%
   mutate(PercentFlagged = round(Flags / Observations*100)) %>%
   arrange(desc(PercentFlagged)) %>% datatable()
```

## Rover Abundance

Indices are converted to the low end of the count bin (1 = 1, 2 = 2, 3 = 10, 4 = 100, 5 = 1000) before comparing to maximum count guideline.

```{r}
# Create flags based on master species list
frs_abund <- frs_raw %>% 
   # convert abundance index into a count
   mutate(MinAbund = case_match(`Abundance index`, 
                                1 ~ 1, 2 ~ 2, 3 ~ 10, 4 ~ 100, 5 ~ 1000),
          MaxAbund = case_match(`Abundance index`, 
                                1 ~ 1, 2 ~ 10, 3 ~ 100, 4 ~ 1000, 5 ~ 10000)) %>% 
   left_join(species,by=c("Scientific name"="scientific name")) %>% 
   mutate(flag = as.factor(ifelse(MinAbund > `max abund`, "yes","no")))
# Calculate percent flagged
rperabundflag <- round(summary(frs_abund$flag)[2] / nrow(frs_abund) *100, 1) 
# Make table of flags
rabund_flag <- frs_abund %>% filter(flag == "yes") %>%
   mutate(Diff = MinAbund - `max abund`) %>% 
   mutate(RelDiff = round(Diff / `max abund`,1)) %>% 
   select(`Site name`,Rep,Name,
          `Scientific name`,`Common name`,`Abundance index`,
          `max abund`, Diff, RelDiff)
```

There were `r rperabundflag`% of rover species flagged on abundance.

```{r, tab.cap="Table 8. Observations from fish rovers flagged on high abundance."}
kable(x=tibble(" "=" ")) # hack caption above datatable
datatable(rabund_flag)
```

```{r, tab.cap="Table 9. Fish species that were most frequently flagged on abundance in rovers. We might want to change the max number on these."}
kable(x=tibble(" "=" ")) # hack caption above datatable
# Summarize abundance flag table by species
rabund_flag %>% 
   group_by(`Common name`,`max abund`) %>% 
   summarize(Flags = length(`Abundance index`), .groups="drop") %>% 
   arrange(desc(Flags)) %>% datatable()
```

```{r, tab.cap="Table 10. Divers who were most frequently flagged on high abundance in rovers. They might be overestimating counts."}
kable(x=tibble(" "=" ")) # hack caption above datatable
# Summarize total number of surveys for each person
frs_namecount <- frs_raw %>% 
   group_by(Name) %>% 
   summarize(Observations = length(Name)) %>% 
   arrange(desc(Observations))
# Standardize flags to a percentage of total fish seen
rabund_flag %>% group_by(Name) %>% 
   summarize(Flags = length(`Abundance index`)) %>% 
   left_join(frs_namecount,by="Name") %>% 
   mutate(PercentFlagged = round(Flags / Observations*100)) %>% 
   arrange(desc(PercentFlagged)) %>% datatable()
```

# Transect Sizing

We use size bins to estimate fish lengths, and we have a maximum and minimum size guideline for each species. An observation is flagged if the small end of the bin is larger than the max size for the species, or the large end of the bin is smaller than the min size for the species.

```{r}
# Create flags based on master species list
fts_size  <- fts_raw %>% 
   # Convert to long format for easier processing
   pivot_longer(cols = X0to5:Xgt150, names_to = "SizeBin",values_to = "Count") %>% 
   # Take out empty observations
   filter(!is.na(Count)) %>% 
   # Rename bins
   mutate(SizeBin = gsub("to","-", gsub("X","",SizeBin))) %>% 
   # Set min and max size of observations based on bin
   mutate(BinMin = case_match(SizeBin, "0-5"~0,"6-10"~6,"11-20"~11,
                              "21-30"~21,"31-40"~31,"41-50"~41,
                              "51-60"~51,"61-70"~61,"71-80"~71,
                              "81-90"~81, "91-100"~91,"101-110"~101,
                              "111-120"~111,"121-130"~121,"131-140"~131,
                              "141-150"~141,"gt150"~151),
          BinMax = case_match(SizeBin, "0-5"~5,"6-10"~10,"11-20"~20,
                              "21-30"~30,"31-40"~40,"41-50"~50,
                              "51-60"~60,"61-70"~70,"71-80"~80,
                              "81-90"~90, "91-100"~100,"101-110"~110,
                              "111-120"~120,"121-130"~130,"131-140"~140,
                              "141-150"~150,"gt150"~151)) %>% 
   # Join to master list to set flags
   left_join(species, by=c("Scientific name"="scientific name")) %>% 
   # If the smallest option is larger than the max size or largest option is smaller than the min size, flag it.
   mutate(flag = as.factor(ifelse(BinMin > `max size` | BinMax < `min size`, "yes","no")))
# Calculate percent flagged
persizeflag <- round(summary(fts_size$flag)[2] / nrow(fts_size) *100, 1) 
# Make table of flags
size_flag <- fts_size %>% filter(flag == "yes") %>%
   select(`Site name`,Rep,Name,
          `Common name`,SizeBin, Count, `min size`, `max size`)
```

There were `r persizeflag`% of transect observations flagged on size.

```{r, tab.cap="Table 11. Observations from fish transects flagged on small or large size."}
kable(x=tibble(" "=" ")) # hack caption above datatable
datatable(size_flag)
```

```{r, tab.cap="Table 12. Fish species that were most frequently flagged on size in transects. We might want to change the size guidelines on these."}
kable(x=tibble(" "=" ")) # hack caption above datatable
# Summarize size flag table by species
size_flag %>%
   group_by(`Common name`,SizeBin,`min size`,`max size`) %>%
   summarize(Flags = length(Count), .groups="drop") %>%
   arrange(desc(Flags)) %>% datatable()
```

```{r, tab.cap="Table 13. Divers who were most frequently flagged on sizes in transects. They might be under- or over-sizing."}
kable(x=tibble(" "=" ")) # hack caption above datatable
# Use total number of transect fish seen from earlier
# Standardize flags to a percentage of total fish seen
size_flag %>% group_by(Name) %>%
   summarize(Flags = length(Count)) %>%
   left_join(fts_namecount,by="Name") %>%
   mutate(PercentFlagged = round(Flags / Observations*100)) %>%
   arrange(desc(PercentFlagged)) %>% datatable()
```

```{r, fig.cap="Figure 1. A closer look at each diver's overall distribution of observed sizes. Look for trends in under- or over-sizing."}
ftsizemelt <- fts_raw %>% 
   # take out masked goby because I'm the only one who counts them
   # it messes up my size distribution
   filter(`Common name`!="masked goby") %>% 
   pivot_longer(cols = c("X0to5":"Xgt150"), 
                names_to = "SizeBin", values_to = "num") %>% 
   mutate(num = replace_na(num, 0))
ftsizemeltSum <- ftsizemelt %>% 
      group_by(Name, SizeBin) %>% 
      summarize(total = sum(num, na.rm=T), .groups="drop_last") %>% 
      mutate(totalrel = total / sum(total)*100,
             SizeBin = factor(gsub("X","",SizeBin), 
                              levels = gsub("X","",
                                            c("X0to5","X6to10","X11to20",
                                              "X21to30","X31to40",
                                              "X41to50","X51to60",
                                              "X61to70","X71to80",
                                              "X81to90","X91to100",
                                              "X101to110","X111to120",
                                              "X121to130","X131to140",
                                              "X141to150","Xgt150")))) %>% 
      ungroup()

ggplot(data=ftsizemeltSum, aes(x=SizeBin,y=totalrel)) +
   geom_bar(aes(alpha=0.5), stat="identity", position="stack", 
            color = "blue") +
   facet_wrap(~Name) +
   guides(alpha="none") +
   ylab("Relative number") +
   theme(axis.text.x = element_text(angle=-90, hjust=0))
```

# Species Richness

## Transects

```{r, tab.cap="Table 14. Summary statistics for number of fish species seen on transects for each diver. Check if they need to learn more ID or may be inflating transect size."}
kable(x=tibble(" "=" ")) # hack caption above datatable
fts_raw %>% 
   group_by(Name, `Fish transect`) %>% 
   summarize(Richness = length(`Common name`), .groups = "drop_last") %>% 
   summarize(n= length(Richness),
             mean=round(mean(Richness),2),
             sd=round(sd(Richness),2),
             min=min(Richness),
             max=max(Richness)) %>% 
   datatable()
```

## Rovers

```{r, tab.cap="Table 15. Summary statistics for number of fish species seen on rovers for each diver. Check if they need to learn more ID."}
kable(x=tibble(" "=" ")) # hack caption above datatable
frs_raw %>% 
   group_by(Name, `Fish rover`) %>% 
   summarize(Richness = length(`Common name`), .groups = "drop_last") %>% 
   summarize(n= length(Richness),
             mean=round(mean(Richness),2),
             sd=round(sd(Richness),2),
             min=min(Richness),
             max=max(Richness)) %>% 
   datatable()
```

# Rare Fish

```{r, tab.cap="Table 16. Fish species with the fewest number of observations (bottom 50). Check for mistaken identity."}
kable(x=tibble(" "=" ")) # hack caption above datatable
rare <- fts_raw %>% 
   group_by(`Common name`) %>% 
   summarize(TotalNum = sum(Total), .groups="drop") %>% 
   arrange(TotalNum) %>% slice_head(n=50) %>% 
   add_column(Rank = c(1:50))
datatable(rare)
```

```{r, tab.cap="Table 17. Identifying information on fish in previous table to help check for mistaken identity."}
kable(x=tibble(" "=" ")) # hack caption above datatable
fts_raw %>% 
   filter(`Common name` %in% rare$`Common name`) %>% 
   left_join(rare, by="Common name") %>% 
   select(`Site name`, Name, Rep, `Common name`, `Scientific name`, Rank) %>% 
   arrange(Rank) %>% datatable()
```

```{r, tab.cap="Table 18. Nassau grouper seen on transects."}
kable(x=tibble(" "=" ")) # hack caption above datatable
fts_raw %>% 
   filter(`Common name`=="Nassau grouper") %>% 
   group_by(`Site name`, Rep, Name) %>% 
   summarize(n = sum(Total), .groups="drop") %>% 
   datatable()
```

```{r, tab.cap="Table 19. Nassau grouper seen on rovers."}
kable(x=tibble(" "=" ")) # hack caption above datatable
frs_raw %>% 
   filter(`Common name`=="Nassau grouper") %>% 
   select(`Site name`, Rep, Name, `Abundance index`) %>% 
   arrange(desc(`Abundance index`)) %>% datatable()
```

```{r, tab.cap="Table 20. Sharks seen on transects."}
kable(x=tibble(" "=" ")) # hack caption above datatable
fts_raw %>% 
   left_join(select(species, `Scientific name`=`scientific name`,
                    Family = family), by="Scientific name") %>% 
   filter(Family=="Carcharhinidae") %>% 
   group_by(`Common name`, `Site name`, Rep, Name) %>% 
   summarize(n = sum(Total), .groups="drop") %>% 
   datatable()
```

```{r, tab.cap="Table 21. Sharks seen on rovers."}
kable(x=tibble(" "=" ")) # hack caption above datatable
frs_raw %>% 
   left_join(select(species, `Scientific name`=`scientific name`,
                    Family = family), by="Scientific name") %>% 
   filter(Family=="Carcharhinidae") %>% 
   select(`Common name`, `Site name`, Rep, Name, 
            `Abundance index`) %>% 
   arrange(desc(`Abundance index`)) %>% 
   datatable()
```

# Diadema

## Overview

```{r, tab.cap="Table 22. Do we have all the Diadema transects? We should have 9 transects each at 33 sites and maybe 1 at Ginsburg's Fringe."}
# Summarize number of diadema surveys by site
das_sitesum <- das_raw %>%
   group_by(`Site name`, Rep) %>% 
   summarize(n=1, .groups="drop_last") %>% 
   summarize(Transects = sum(n))
# High level look at number of reps
das_sitesum %>% 
   group_by(Transects) %>% 
   summarize(Sites = length(`Site name`)) %>% kable()
```

```{r, tab.cap="Table 23. Sites with fewer than 9 Diadema transects."}
kable(x=tibble(" "=" ")) # hack caption above datatable
# See the sitewise list
das_sitesum %>% filter(Transects<9) %>% arrange(Transects) %>% datatable()
```

## Diadema Abundance

```{r, tab.cap="Table 24. Summary statistics for number of Diadema seen on transects at each site. Most sites usually have none."}
kable(x=tibble(" "=" ")) # hack caption above datatable
# A 0 test size means diadema were looked for and not seen
# A NA test size means diadema were seen but not sized
das_raw %>% 
   # Convert to presence/absence: set 0 size to 0 and rest to 1
   mutate(PA = ifelse(`Test size cm`==0 & !is.na(`Test size cm`), 
                      0, 1)) %>% 
   group_by(`Site name`, Rep) %>% 
   # Add to get count per transect
   summarize(Count = sum(PA), .groups="drop") %>% 
   # Summarize count by site
   group_by(`Site name`) %>% 
   summarize(n = length(Count), mean = round(mean(Count),2), 
             sd = round(sd(Count),2),
             min = min(Count), max = max(Count), 
             .groups="drop") %>% 
   arrange(desc(mean)) %>% datatable()
```

## Diadema Sizing

```{r, tab.cap="Table 25. Diadema with a test size estimated larger than 15 cm."}
kable(x=tibble(" "=" ")) # hack caption above datatable
das_raw %>% 
   # Take out zeros
   filter(`Test size cm`>=15) %>% datatable()
# Count number of NAs
unsized <- das_raw %>% filter(is.na(`Test size cm`)) %>% nrow()
```

There were `r unsized` Diadema observed but not sized.

```{r, fig.cap="Figure 2. Size frequency distribution of observed and sized Diadema, colored by site."}
das_raw %>% 
   # Take out zeros
   filter(`Test size cm`>0) %>% 
   # Plot size frequency colored by site
   ggplot(data=., aes(x=`Test size cm`, fill=`Site name`)) +
   geom_histogram(binwidth=1) +
   scale_x_continuous(breaks=seq(0,15,1)) +
   theme_minimal()
```

```{r, tab.cap="Table 26. All Diadema observed."}
kable(x=tibble(" "=" ")) # hack caption above datatable
das_raw %>% 
   # Take out zeros but not NAs
   filter(`Test size cm`!=0 | is.na(`Test size cm`)) %>% 
   select(-Notes,-`Fish transect`) %>% 
   datatable()
```

# Changes Made

-   Unsized Diadema were added into counts with a size of NA.

-   Rover BTY 1 O'Neill red hind index incorrectly entered as 3, changed to 2.

-   Rover SVN 2 O'Neill puddingwife index of 3 changed to wrasse species.

-   Rover MGN 1 Hobbs longfin damsel index of 4 thought to be accidental tally and changed to 2. Only 1 other person saw 1 of this species at the whole site.

-   Rover SVN 3 Hobbs tallied instead of using index and some were entered incorrectly so these were all checked and changed to the proper index.

-   Rover SSJ 1 Kadison barred hamlet index of 3 changed to 2 since no one else saw more than 1.

-   Rover GRP 2 Hollander cero index of 3 changed to 2, thought to be accidental tally.

-   Rover MTS 2 Hollander sharpnose puffer index of 3 thought to be accidental tally and changed to 2.

-   Rover MGN 2 O'Neill cocoa damselfish tallies were faded so not entered; added.

-   Transect BTY 2 Hobbs trumpetfish 6-10cm changed to 11-20cm. All her fish appear on the small side; this should be mentioned for in-water correction on future surveys.

-   Transect BIT 2 Maxim & BTY 8 O'Neill white margate thought to be unlikely and changed to grunt species.

-   Transect BPT 4 Parr yellowbelly hamlets 21-30cm changed to 11-20cm.

-   Transect CBS 8 Hollander trumpetfish each bumped up 1 size bin.

-   Transect LBP 5 Arrington rock beauty tally in \>40cm was erased on datasheet so we deleted. We added one for him in 11-20cm on LBP 3 where there was no tally but it was obviously seen.

-   Transect CBS 7 Arrington blue hamlet changed to indigo hamlet.

-   Transect CBS 7 Arrington night sergeant thought to be entry mistake and changed to sergeant major.

-   Transect BTY 1 Shelby 0-5 cm yellowtail parrotfish changed to parrotfish species.

-   Transect SRD 1 Dade bantam bass only written because of pressure from TBS; deleted.

-   Transect SVN 1 Shelby striped grunt changed to grunt species.

-   Transect CST 1 Hollander painted wrasse changed to wrasse species.
