---
title: "TCRMP_2023_Benthic_QAQC.qmd"
author: "Sarah Heidmann"
format: html
editor: visual
---

### Sarah Heidmann

### 2024-09-24

QAQC for TCRMP coral health and algae height data for 2023 and 2024 Post Bleaching (PBL). Suggestions for additions and improvement are welcomed.

Luckily, the data entry system catches a lot of easy mistakes like mispellings. We just have to go through the mistakes it might not catch. When we catch some that are fixable, change online and export again, or if the online database has already been cleared, change directly in the data.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE)
library(tidyverse, quietly = TRUE)
library(knitr)
options(knitr.kable.NA = '')
library(DT)

# Read qaqc'd data for fish and diadema
chs_raw_23 <- read_csv("data/2023/TCRMP_CoralHealth_2023_webexport.csv")
chs_raw_24 <- read_csv("data/2023/TCRMP_CoralHealth_2024PBL_webexport.csv")
ahs_raw <- read_csv("data/2023/TCRMP_AlgaeHeights_2023_webexport.csv")

# Species master for QAQC numbers
#species <- read_csv("data/fish_species_list_2022.csv")

```

# Overview

How big are the datasets?

```{r}
corals23 <- length(unique(chs_raw_23$`Transect coral`))
corals24 <- length(unique(chs_raw_24$`Transect coral`))
```

Coral health 2023 has `r corals23` corals. Coral health 2024PBL has `r corals24` corals.

Do we have all the coral transects in 2023? (should have 6 transects each at 33 sites)

```{r}
# Summarize number of transects by site
chs_sitesum_23 <- chs_raw_23 %>% 
   group_by(`Site name`,Rep) %>% 
   summarize(n=1, .groups="drop_last") %>% 
   summarize(Transects = sum(n))
# High level look at number of reps
chs_sitesum_23 %>% 
   group_by(Transects) %>% 
   summarize(Sites = length(`Site name`)) %>% kable()
```

Those with less than 6 transects are below.

```{r}
# See the sitewise list
chs_sitesum_23 %>% filter(Transects<6) %>% arrange(Transects) %>% datatable()
```

Do we have all the coral transects in 2024PBL? (should have 6 transects each at 33 sites)

```{r}
# Summarize number of transects by site
chs_sitesum_24 <- chs_raw_24 %>% 
   group_by(`Site name`,Rep) %>% 
   summarize(n=1, .groups="drop_last") %>% 
   summarize(Transects = sum(n))
# High level look at number of reps
chs_sitesum_24 %>% 
   group_by(Transects) %>% 
   summarize(Sites = length(`Site name`)) %>% kable()
```

Those with less than 6 transects are below.

```{r}
# See the sitewise list
chs_sitesum_24 %>% filter(Transects<6) %>% arrange(Transects) %>% datatable()
```

Do we have all the algae transects in 2023? (should have 6 transects each at 33 sites)

```{r}
# Summarize number of transects by site
ahs_sitesum <- ahs_raw %>% 
   group_by(`Site name`,Rep) %>% 
   summarize(n=1, .groups="drop_last") %>% 
   summarize(Transects = sum(n))
# High level look at number of reps
ahs_sitesum %>% 
   group_by(Transects) %>% 
   summarize(Sites = length(`Site name`)) %>% kable()
```

Those with less than 6 transects are below.

```{r}
# See the sitewise list
ahs_sitesum %>% filter(Transects<6) %>% arrange(Transects) %>% datatable()
```

## Date ranges

```{r}
# Date ranges for each survey type
ch_range_23 <- range(chs_raw_23$`Date completed`)
ch_range_24 <- range(chs_raw_24$`Date completed`)
ah_range <- range(ahs_raw$`Date completed`)
```

Coral health 2023 took place between dates: `r ch_range_23`

Coral health 2024PBL took place between dates: `r ch_range_24`

Algae heights took place between dates: `r ah_range`
