---
title: "TCRMP 2003-2015 Fish Rovers QAQC"
author: "Sarah Heidmann"
format: html
editor: visual
toc: true
---

### 2026-03-27

QAQC for TCRMP fish rover data for 2003-2015, entered in 2025-2026. Suggestions for additions and improvement are welcomed.

Luckily, the data entry system catches a lot of easy mistakes like mispellings. We just have to go through the mistakes it might not catch. When we catch some that are fixable, change online and export again, or if the online database has already been cleared, change directly in the data.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE)
library(tidyverse, quietly = TRUE)
library(knitr)
options(knitr.kable.NA = '')
library(DT)

# Read qaqc'd data
frs_raw <- read_csv("data/TCRMP_FishRovers_2003_2015_webexport.csv",
                    col_types = cols(`Date completed`=col_date(format="%m/%d/%Y"))) %>% 
   # add year for easier summarizing
   mutate(SampleYear = format(`Date completed`, "%Y"))

# Species master for QAQC numbers
species <- read_csv("data/fishlists/FishMetadata_202602.csv")

```

# Overview

How big is the dataset?

Rovers have `r nrow(frs_raw)` rover-species.

```{r, tab.cap="Table 1. Do we have all the fish rovers? A normal year should have 3 reps each at 33 sites and maybe 1 at Ginsburg's Fringe."}
# Summarize number of rovers by site and year
frs_sitesum <- frs_raw %>% 
   group_by(SampleYear,`Site name`,Rep) %>% 
   summarize(n=1, .groups="drop_last") %>% 
   summarize(Rovers = sum(n), .groups="drop")
# High level look at number of reps
frs_sitesum %>% 
   group_by(Rovers) %>% 
   summarize("Year-Sites" = length(`Site name`)) %>% kable()
```

```{r, tab.cap="Table 2. Year-sites with less or more than 3 fish rovers."}
# See the sitewise list
frs_sitesum %>% filter(Rovers !=3) %>% arrange(Rovers) %>% datatable()
```

## Date ranges

```{r}
# Date range
fr_range <- range(frs_raw$`Date completed`)
```

Fish rovers took place between dates: `r fr_range`

## Notes

Are there any notes we should know about?

```{r}
frs_raw %>% filter(!is.na(Notes)) %>% 
   pull(Notes) %>% 
   unique()
```

# Abundance

An observation is flagged if the number of fish observed is greater than the maximum number guideline.

## Rover Abundance

Indices are converted to the low end of the count bin (1 = 1, 2 = 2, 3 = 10, 4 = 100, 5 = 1000) before comparing to maximum count guideline.

```{r}
# Create flags based on master species list
frs_abund <- frs_raw %>% 
   # convert abundance index into a count
   mutate(MinAbund = case_match(`Abundance index`, 
                                1 ~ 1, 2 ~ 2, 3 ~ 10, 4 ~ 100, 5 ~ 1000),
          MaxAbund = case_match(`Abundance index`, 
                                1 ~ 1, 2 ~ 10, 3 ~ 100, 4 ~ 1000, 5 ~ 10000)) %>% 
   left_join(species,by=c("Scientific name"="ScientificName")) %>% 
   mutate(flag = as.factor(ifelse(MinAbund > MaxNum, "yes","no")))
# Calculate percent flagged
rperabundflag <- round(summary(frs_abund$flag)[2] / nrow(frs_abund) *100, 1) 
# Make table of flags
rabund_flag <- frs_abund %>% filter(flag == "yes") %>%
   mutate(Diff = MinAbund - MaxNum) %>% 
   mutate(RelDiff = round(Diff / MaxNum,1)) %>% 
   select(SampleYear, `Site name`,Rep,Name,
          `Scientific name`,`Common name`,`Abundance index`,
          MaxNum, Diff, RelDiff)
```

There were `r rperabundflag`% of rover species flagged on abundance.

```{r, tab.cap="Table 3. Observations from fish rovers flagged on high abundance."}
kable(x=tibble(" "=" ")) # hack caption above datatable
datatable(rabund_flag)
```

```{r, tab.cap="Table 4. Fish species that were most frequently flagged on abundance in rovers. We might want to change the max number on these."}
kable(x=tibble(" "=" ")) # hack caption above datatable
# Summarize abundance flag table by species
rabund_flag %>% 
   group_by(`Common name`,MaxNum) %>% 
   summarize(Flags = length(`Abundance index`), .groups="drop") %>% 
   arrange(desc(Flags)) %>% datatable()
```

```{r, tab.cap="Table 5. Divers who were most frequently flagged on high abundance in rovers. They might be overestimating counts."}
kable(x=tibble(" "=" ")) # hack caption above datatable
# Summarize total number of surveys for each person
frs_namecount <- frs_raw %>% 
   group_by(Name) %>% 
   summarize(Observations = length(Name)) %>% 
   arrange(desc(Observations))
# Standardize flags to a percentage of total fish seen
rabund_flag %>% group_by(Name) %>% 
   summarize(Flags = length(`Abundance index`)) %>% 
   left_join(frs_namecount,by="Name") %>% 
   mutate(PercentFlagged = round(Flags / Observations*100)) %>% 
   arrange(desc(PercentFlagged)) %>% datatable()
```

# Species Richness

## Rovers

```{r, tab.cap="Table 6. Summary statistics for number of fish species seen on rovers for each diver. Check if they need to learn more ID (jk too late)."}
kable(x=tibble(" "=" ")) # hack caption above datatable
frs_raw %>% 
   group_by(Name, `Fish rover`) %>% 
   summarize(Richness = length(`Common name`), .groups = "drop_last") %>% 
   summarize(n= length(Richness),
             mean=round(mean(Richness),2),
             sd=round(sd(Richness),2),
             min=min(Richness),
             max=max(Richness)) %>% 
   datatable()
```

# Rare Fish

```{r, tab.cap="Table 7. Fish species with the fewest number of observations (bottom 50). Check for mistaken identity."}
kable(x=tibble(" "=" ")) # hack caption above datatable
rare <- frs_abund %>% 
   group_by(`Common name`) %>% 
   summarize(TotalNum = sum(MinAbund), .groups="drop") %>% 
   arrange(TotalNum) %>% slice_head(n=50) %>% 
   add_column(Rank = c(1:50))
datatable(rare)
```

```{r, tab.cap="Table 8. Identifying information on fish in previous table to help check for mistaken identity."}
kable(x=tibble(" "=" ")) # hack caption above datatable
frs_raw %>% 
   filter(`Common name` %in% rare$`Common name`) %>% 
   left_join(rare, by="Common name") %>% 
   select(SampleYear,`Site name`, Name, Rep, 
          `Common name`, `Scientific name`, Rank) %>% 
   arrange(Rank) %>% datatable()
```

```{r, tab.cap="Table 9. Nassau grouper seen on rovers."}
kable(x=tibble(" "=" ")) # hack caption above datatable
frs_raw %>% 
   filter(`Common name`=="Nassau grouper") %>% 
   select(SampleYear,`Site name`, Rep, Name, `Abundance index`) %>% 
   arrange(desc(`Abundance index`)) %>% datatable()
```

```{r, tab.cap="Table 10. Sharks seen on rovers."}
kable(x=tibble(" "=" ")) # hack caption above datatable
frs_raw %>% 
   left_join(select(species, `Scientific name`=ScientificName, Family), 
             by="Scientific name") %>% 
   filter(Family=="Carcharhinidae") %>% 
   select(`Common name`, SampleYear, `Site name`, Rep, Name, 
            `Abundance index`) %>% 
   arrange(desc(`Abundance index`)) %>% 
   datatable()
```

# Changes Made

-   There are lots of flags but I don't know what reefs looked like 20 years ago and have little reason to change the data.
-   I did change 1000 dash goby in 2008 at Seahorse 1 to masked goby
